---
title: "Getting started with parallax"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Getting started with parallax}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
library(parallax)
```

`parallax` wraps point cloud workflows in platform-aware R classes and uses a
Rust backend for the operations that are likely to become expensive on real
data. The installed demo files are intentionally tiny, but they mirror the same
flow you would use for LAS/LAZ data.

## Demo files

```{r}
aerial_path <- system.file("extdata", "aerial-demo.csv", package = "parallax")
terrestrial_path <- system.file("extdata", "terrestrial-demo.csv", package = "parallax")

aerial <- px_read(aerial_path, platform = "aerial")
terrestrial <- px_read(terrestrial_path, platform = "terrestrial")

class(aerial)
class(terrestrial)
```

## Aerial workflow

Start with filtering and ground classification.

```{r}
filtered <- px_filter_statistical(aerial, k = 6, std_ratio = 1.5)
ground <- px_classify_ground(filtered)

c(
  n_points = nrow(filtered$xyz),
  n_ground = ground$n_ground,
  n_nonground = ground$n_nonground
)
```

The same cloud can feed tree segmentation and summary metrics.

```{r}
trees <- px_segment_trees(filtered, min_height = 2.0, crown_threshold = 0.75)
metrics <- px_metrics(filtered)

trees$n_segments
metrics
```

If you want gridded outputs, assign a simple ground/non-ground classification
and build canopy products.

```{r}
filtered$classification <- ifelse(ground$ground_mask, 2L, 1L)

chm <- px_canopy_height_model(filtered, resolution = 1.0)
density <- px_density_map(filtered, resolution = 1.0)

dim(chm$chm)
dim(density$density)
```

## Terrestrial workflow

Terrestrial data often leads with local geometry. Here the small demo scan has
a dominant plane and enough points for local normal estimation.

```{r}
scan_with_normals <- px_estimate_normals(terrestrial, k_neighbors = 4L)
planes <- px_detect_planes(
  scan_with_normals,
  distance_threshold = 0.08,
  min_points = 6L,
  num_iterations = 100L
)

length(planes)
if (length(planes) > 0) planes[[1]]$equation
```

## Where to go next

- Use `px_read()` on LAS/LAZ files for real data.
- Use `px_icp()` and `px_align_scans()` for scan registration.
- Use `px_poisson_reconstruct()` after `px_estimate_normals()` for mesh-style
  outputs.
- Convert to and from `lidR` objects with `to_lidr()` and `from_lidr()`.
