---
title: "Platform-aware workflows"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Platform-aware workflows}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
library(parallax)
```

The package API stays consistent across platforms, but defaults and typical
follow-on tasks differ between aerial and terrestrial data.

## Constructors and metadata

```{r}
aerial <- px_read(
  system.file("extdata", "aerial-demo.csv", package = "parallax"),
  platform = "aerial"
)

scan <- px_read(
  system.file("extdata", "terrestrial-demo.csv", package = "parallax"),
  platform = "terrestrial"
)

aerial$metadata$platform
scan$metadata$platform
```

## The same function, different context

Ground classification runs through one function for both clouds, but the
meaning of the result is tied to the acquisition platform.

```{r}
aerial_ground <- px_classify_ground(aerial)
scan_ground <- px_classify_ground(scan)

c(aerial_ground = aerial_ground$n_ground, terrestrial_ground = scan_ground$n_ground)
```

Tree segmentation makes sense for the aerial scene, while plane detection is a
better fit for the terrestrial scan.

```{r}
trees <- px_segment_trees(aerial, min_height = 2.0, crown_threshold = 0.75)
planes <- px_detect_planes(
  px_estimate_normals(scan, k_neighbors = 4L),
  distance_threshold = 0.08,
  min_points = 6L,
  num_iterations = 100L
)

trees$n_segments
length(planes)
```

## Registration and reconstruction

The installed registration example ships as a source/target pair.

```{r}
source <- px_read(
  system.file("extdata", "registration-source.csv", package = "parallax"),
  platform = "terrestrial"
)
target <- px_read(
  system.file("extdata", "registration-target.csv", package = "parallax"),
  platform = "terrestrial"
)

icp <- px_icp(source, target, max_correspondence_distance = 2.0)
icp$transformation
```

Once a cloud has normals, it can move into reconstruction-oriented functions.

```{r}
target_with_normals <- px_estimate_normals(target, k_neighbors = 4L)
mesh <- px_poisson_reconstruct(target_with_normals, depth = 4L)

c(vertices = mesh$n_vertices, faces = mesh$n_faces)
```

## Takeaway

The package surface is intentionally repetitive: read, filter, segment,
register, and reconstruct with the same `px_*` verbs, and let the cloud class
carry the platform context.
